hormones

### 2020-01-15 analysis/07_hormones.Rmd
more figurs

@@

@@

plotcorrelations <- function(i,j, myxlab, mylab, mysubtitle){
      p <- ggplot(hormoneswide, aes(x = i, y = j),na.rm = TRUE) +
       geom_point(aes(color = treatment), na.rm = TRUE) +
       geom_smooth( method = lm, color = "darkgrey", na.rm = TRUE ) +
       theme_B3() +
      #facet_wrap(~sex, scales = "free") +
        labs(x = myxlab, y = mylab, subtitle = mysubtitle) +
        theme(legend.position = "none")

@@

plot_grid(p1,p3 + theme(legend.position = "right"), align = "hv", rel_widths = c(0.4,0.6))

@@

>>>>

theme_B3 <- function () { 
  theme_classic(base_size = 11) +
    theme(
      panel.grid.major  = element_blank(),  # remove major gridlines
      panel.grid.minor  = element_blank(),  # remove minor gridlines
      plot.title = element_text(hjust = 0.5, face = "bold"), # center & bold 
      axis.text.x = element_text(angle = 45, hjust = 1)
    )
}

@@

## prolactin 2

```{r PRL2}
hormones %>%
  filter(hormone == "prolactin") %>%
  group_by(sex) %>%
  summarise(m = mean(plasma_conc),
            median = median(plasma_conc))

df <- hormones %>%
  filter(hormone == "prolactin",
         study == "characterization")
p2 <-  ggplot(df, aes(x = treatment, y = plasma_conc, fill = treatment, color = sex)) +
  geom_boxplot(aes(color = sex)) +
  theme_B3() +
  labs(y = NULL, x = NULL, subtitle = " ") +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = sexcolors) +
  scale_y_continuous(limits = c(0,105)) +
    geom_text(aes(label = "B", x = 1, y = 100, fontface = "plain"), size = 3) +
   geom_text(aes(label = "A", x = 2, y = 100), size = 3) +
geom_text(aes(label = "A", x = 3, y = 100), size = 3) +
geom_text(aes(label = "B", x = 4, y = 100), size = 3) +
  geom_text(aes(label = "C", x = 5, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 6, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 7, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 8, y = 100), size = 3) +
  geom_text(aes(label = "D", x = 9, y = 100), size = 3) +
  geom_text(aes(label = '*', x = 1, y = 105), size = 5) +
   geom_text(aes(label = '*', x = 2, y = 105), size = 5) 
  p2

p3 <- hormones %>%
  filter(hormone == "prolactin",
         treatment %in% c(controlsremoval, levelsremoval )) %>% 
  ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot(aes(color = sex)) +
  theme_B3() +
  scale_y_continuous(limits = c(0,110)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs( x = NULL, y = "concentration (ng/mL", subtitle = " ") +
  scale_color_manual(values = sexcolors) +
   geom_text(aes(label = '*', x = 2, y = 105), size = 5) 
p3

p4 <- hormones %>%
  filter(hormone == "prolactin",
         treatment %in% c(controlstiming, levelstiming )) %>% 
ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot(aes(color = sex)) +
  theme_B3() +
  scale_y_continuous(limits = c(0,110)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(y = NULL, x = NULL, subtitle = " ") +
  scale_color_manual(values = sexcolors) 
p4

p12 <- plot_grid(p1, p2, rel_widths = c(0.15,0.8), labels = "auto", label_size = 8)
p34 <- plot_grid(p3,p4, nrow = 1, rel_widths = c(0.55,0.45), labels = c("c", "d"), label_size = 8)

plot_grid(p12, p34, nrow = 2)
```

## correlations

@@

plotcorrelations <- function(i,j, myxlab, mylab, mysubtitle){
      p <- ggplot(hormoneswide, aes(x = i, y = j),na.rm = TRUE) +
       geom_point(aes(color = treatment), na.rm = TRUE) +
       geom_smooth( method = lm, color = "darkgrey", na.rm = TRUE ) +
       theme_B3() +
      #facet_wrap(~sex, scales = "free") +
        labs(x = myxlab, y = mylab, subtitle = mysubtitle) +
        theme(legend.position = "none") +
        scale_color_manual(values = colorscharmaip2)

@@

plot_grid(p1 + theme(axis.text.x = element_text(angle = 0)),p3+ theme(axis.text.x = element_text(angle = 0))  + theme(legend.position = "right"), align = "hv", rel_widths = c(0.4,0.6))

@@

## progesterone

```{r progesterone}

hormones %>%
  filter(hormone == "progesterone") %>%
  group_by(sex) %>%
  summarise(mean = mean(plasma_conc),
            median = median(plasma_conc))

p1 <- hormones %>%
  filter(hormone == "progesterone",
         study == "characterization") %>%
  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
  geom_boxplot() +
  theme_B3() +
  labs(y = "concentration (ng/mL)", subtitle = "Progesterone" , x = NULL) +
  theme(legend.position = "none")   +
  scale_y_continuous(limits = c(0,4.5)) +
  scale_fill_manual(values = sexcolors)

df <- hormones %>%
  filter(hormone == "progesterone",
         study == "characterization")

p2 <-  ggplot(df, aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  labs(y = NULL, x = NULL, subtitle = " ") +
  scale_y_continuous(limits = c(0,4.5)) +
  theme(legend.position = "none")

p3 <- hormones %>%
  filter(hormone == "progesterone",
         treatment %in% c(controlsremoval, levelsremoval )) %>% 
ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  scale_y_continuous(limits = c(0,4.5)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none") +
  labs( x = NULL, y = "concentration (ng/mL")

p4 <- hormones %>%
  filter(hormone == "progesterone",
         treatment %in% c(controlstiming, levelstiming )) %>% 
ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  scale_y_continuous(limits = c(0,4.5)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none") +
  labs(y = NULL, x = NULL)

progesterone <- plot_grid(p1, p2, p3 + labs(y = NULL), p4, nrow = 1, rel_widths =   c(0.175,0.4,0.35,0.25))
progesterone
```

## estradiol and testosterone

```{r estradiol}

hormones %>%
  filter(hormone == "estradiol") %>%
  summarise(mean = mean(plasma_conc),
            median = median(plasma_conc))

p5 <- hormones %>%
  filter(hormone == "estradiol",
         study == "characterization") %>%
  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
  geom_boxplot() +
  theme_B3() +
  labs(y = "concentration (ng/mL)", subtitle = "Estradiol" , x = NULL) +
  theme(legend.position = "none")  +
    scale_y_continuous(limits = c(0,1)) +
  scale_fill_manual(values = sexcolors)

df <- hormones %>%
  filter(hormone == "estradiol",
         study == "characterization")

p6 <-  ggplot(df, aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  labs(y = NULL, x = NULL, subtitle = " ") +
    scale_y_continuous(limits = c(0,1)) +
  theme(legend.position = "none")

p7 <- hormones %>%
  filter(hormone == "estradiol",
         treatment %in% c(controlsremoval, levelsremoval )) %>% 
ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  scale_y_continuous(limits = c(0,1)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none") +
  labs( x = NULL, y = "concentration (ng/mL")

p8 <- hormones %>%
  filter(hormone == "estradiol",
         treatment %in% c(controlstiming, levelstiming )) %>% 
ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  scale_y_continuous(limits = c(0,1)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none") +
  labs(y = NULL, x = NULL)

estradiol <- plot_grid(p5, p6, p7 + labs(y = NULL), p8, nrow = 1, rel_widths =   c(0.175,0.4,0.35,0.25))
estradiol
```

```{r testosterone}

hormones %>%
  filter(hormone == "testosterone") %>%
  summarise(mean = mean(plasma_conc),
            median = median(plasma_conc))

p9 <- hormones %>%
  filter(hormone == "testosterone",
         study == "characterization") %>%
  ggplot(aes(x = sex, y = plasma_conc, fill = sex)) +
  geom_boxplot() +
  theme_B3() +
  labs(y = "concentration (ng/mL)", subtitle = "Testosterone" , x = NULL) +
  theme(legend.position = "none")  +
    scale_y_continuous(limits = c(0,8.5)) +
  scale_fill_manual(values = sexcolors)

df <- hormones %>%
  filter(hormone == "testosterone",
         study == "characterization")

p10 <-  ggplot(df, aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  labs(y = NULL, x = NULL, subtitle = " ") +
    scale_y_continuous(limits = c(0,8.5)) +
  theme(legend.position = "none")

p11 <- hormones %>%
  filter(hormone == "testosterone",
         treatment %in% c(controlsremoval, levelsremoval )) %>% 
ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  scale_y_continuous(limits = c(0,8.5)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none") +
  labs( x = NULL, y = "concentration (ng/mL")

p12 <- hormones %>%
  filter(hormone == "testosterone",
         treatment %in% c(controlstiming, levelstiming )) %>% 
ggplot(aes(x = treatment, y = plasma_conc, fill = treatment)) +
  geom_boxplot() +
  theme_B3() +
  scale_y_continuous(limits = c(0,8.5)) +
  scale_fill_manual(values = colorscharmaip) +
  theme(legend.position = "none") +
  labs(y = NULL, x = NULL)
p12

testosterone <- plot_grid(p9, p10, p11 + labs(y = NULL), p12, nrow = 1, rel_widths =   c(0.175,0.4,0.35,0.25))
testosterone
```

```{r supplefig, fig.height=8}
plot_grid(p1, p2, p3 + labs(y = NULL), p4, p5, p6, p7 + labs(y = NULL), p8, 
          p9, p10, p11 + labs(y = NULL), p12, nrow = 3, rel_widths =   c(0.175,0.4,0.35,0.25), align = "hv",
          label_size = 8, labels = c("a", " ", " " , " ",
                                     "b", " ", " " , " ",
                                     "c", " ", " " , " "))
